---
title: "Evaluating MCMC"
author: "Alex Cope"
date: '2023-05-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(cowplot)
library(ggpubr)


getPosteriorTable <- function(model.fit.directory,file.name="adaptive_mcmc.Rda")
{
  if (file.exists(file.path(model.fit.directory,file.name)))
  {
    load(file.path(model.fit.directory,file.name))
    df <- adapt.fit$Summary1 %>% 
        as.data.frame() %>%
        rownames_to_column("Parameter")
      write_tsv(df,file.path(model.fit.directory,"posterior_means.tsv"))
  }
}

getDIC <- function(model.fit)
{
  load(model.fit)
  if (!is.na(fit$DIC2[3]))
  {
    df <- data.frame(Mean.LL = mean(fit$Monitor[,"LL"]),Mean.Dev = fit$DIC2[1],Eff.Param=fit$DIC2[2],DIC=fit$DIC2[3],Stationary="Yes")
  } else{
    df <- data.frame(Mean.LL = mean(fit$Monitor[,"LL"]),Mean.Dev = fit$DIC1[1],Eff.Param=fit$DIC1[2],DIC=fit$DIC1[3],Stationary="No")
  }
  return(df)
}

getAdaptDIC <- function(model.fit)
{
  load(model.fit)
  if (!is.na(adapt.fit$DIC2[3]))
  {
    df <- data.frame(Mean.LL = mean(adapt.fit$Monitor[,"LL"]),Mean.Dev = adapt.fit$DIC2[1],Eff.Param=adapt.fit$DIC2[2],DIC=adapt.fit$DIC2[3],Stationary="Yes")
  } else{
    df <- data.frame(Mean.LL = mean(adapt.fit$Monitor[,"LL"]),Mean.Dev = adapt.fit$DIC1[1],Eff.Param=adapt.fit$DIC1[2],DIC=adapt.fit$DIC1[3],Stationary="No")
  }
  return(df)
}


```


# Simulate data with $H$, $Sigma$ the same across all genes -- 11 species

```{r}
truth <- read_tsv("Simulated_data/true_parameters_used_for_parameters_same_sim_100_loci.tsv")
truth[,"H_Y"] <- log(truth[,"H_Y"])
truth[,"H_X"] <- log(truth[,"H_X"])
truth[,"Sigma_Y"] <- log(truth[,"Sigma_Y"])
truth[,"Sigma_X"] <- log(truth[,"Sigma_X"])

truth <- truth %>%
  mutate(Simulation=paste0("Simulated_data_",1:nrow(truth))) %>%
  pivot_longer(-Simulation,names_to="Parameter",values_to="Truth")
  
```


```{r}
directory <- "Results/Test_DIC/Truth_C_eq_0/100_loci/C_eq_0_fit/"
true.models <- list.files(directory,pattern="Simulated_data_[0-9]+",full.names = T)
lapply(true.models,getPosteriorTable)

```


```{r}
directory <- "Results/Test_DIC/Truth_C_eq_0/100_loci/C_eq_0_fit/"
model.fits <- list.files(directory,pattern="posterior_means.tsv", recursive = T)
names(model.fits) <- str_match(model.fits,"Simulated_data_[0-9]{0,2}")

post.mean.df <- lapply(model.fits,function(x)
  {
    read_tsv(file.path(directory,x),col_types =cols()) 
}
)%>% bind_rows(.id="Simulation")

post.mean.df <- post.mean.df %>%
  left_join(truth,by=c("Simulation","Parameter"))
```


```{r fig.width=7,fig.height=7}
post.mean.df.shared <- post.mean.df %>%
  filter(Parameter %in% c("H_Y","H_X","Q","Sigma_Y","Sigma_X"))
post.mean.df.shared <- post.mean.df.shared %>%
  mutate(Parameter = case_when(
    Parameter == "H_X" ~ "alpha[P]",
    Parameter == "H_Y" ~ "alpha[R]",
    Parameter == "Q" ~ "a[R]",
    Parameter == "Sigma_X" ~ "tau[P]",
    Parameter == "Sigma_Y" ~ "tau[R]",
    T ~ Parameter
  ))

p <- ggplot(post.mean.df.shared,aes(x=Truth,y=Mean)) +
  geom_point() +
  stat_cor(method="spearman",label.sep="\n") +
  geom_errorbar(aes(ymin=LB,ymax=UB)) +
  theme_cowplot()+
  ylab("Posterior Mean Estimate") +
  geom_abline(intercept=0,slope=1,linetype="dashed") +
  theme(aspect.ratio=1) +
  facet_wrap(~Parameter,scales="free",labeller = "label_parsed")
p

ggsave2("simulation_results_11_species_100_loci.pdf",p,width=8,height=8)

```


# Model comparisons

## Protein-driven model truth

```{r}
true.models <- list.files("Results/Test_DIC/Truth_C_eq_0/100_loci/C_eq_0_fit/",pattern="rwm_mcmc_1.Rda",recursive = T,full.names = T)
wrong.models <- list.files("Results/Test_DIC/Truth_C_eq_0/100_loci/C_neq_0_reverse_trait_fit/",pattern="rwm_mcmc_1.Rda",recursive = T,full.names = T)

true.data.set <- str_extract(true.models,"Simulated_data_[0-9]+")
wrong.data.set <- str_extract(wrong.models,"Simulated_data_[0-9]+")

names(true.models) <- true.data.set
names(wrong.models) <- wrong.data.set

dic.truth <- lapply(true.models,getDIC) %>% bind_rows(.id="Data")
dic.wrong <- lapply(wrong.models,getDIC) %>% bind_rows(.id="Data")

dic.df <- dic.truth %>%
  inner_join(dic.wrong,by="Data",suffix=c("_True","_Wrong")) %>%
  mutate(Delta.DIC=DIC_Wrong - DIC_True)

```

```{r fig.width=5}

dic.hist <- ggplot(dic.df,aes(x=Delta.DIC)) +
  geom_bar() +
  scale_x_binned() +
  theme_cowplot() +
  xlab(expression(Delta*"DIC")) +
  ylab("Count")
dic.hist

dic.df.stationary <- dic.df %>%
  filter(Stationary_True == "Yes" & Stationary_Wrong == "Yes")

dic.hist.stat <- ggplot(dic.df.stationary,aes(x=Delta.DIC)) +
  geom_bar() +
  scale_x_binned() +
  xlab(expression(Delta*"DIC")) +
  theme_cowplot() +
  ggtitle("Truth: protein-driven model")
dic.hist.stat


dic.cutoff <- c(0,2,10,100,200,300)
names(dic.cutoff) <- dic.cutoff
delta.dic.score <- purrr::map_df(dic.cutoff,function(cutoff)
  {
   return((dic.df.stationary %>% filter(Delta.DIC > cutoff) %>% nrow()) / dic.df.stationary %>% filter(abs(Delta.DIC) > cutoff) %>% nrow()) 
}) %>% pivot_longer(everything(),names_to="Cutoff",values_to="Fraction_correct") %>%
  mutate(Cutoff = as.numeric(Cutoff))

dic.acc <- ggplot(delta.dic.score,aes(x=Cutoff,y=Fraction_correct)) +
  geom_point() +
  theme_cowplot() +
  xlab(expression(Delta*"DIC cutoff")) +
  ylab("Fraction of simulations correct model selected")
dic.acc

prot.driven.dic.perf <- (dic.hist.stat | dic.acc)



```

## mRNA-driven model is truth

```{r}
true.models <- list.files("Results/Test_DIC/Truth_C_neq_0/100_loci/C_neq_0_fit/",pattern="rwm_mcmc_1.Rda",recursive = T,full.names = T)
wrong.models <- list.files("Results/Test_DIC/Truth_C_neq_0/100_loci/C_eq_0_reverse_trait_fit/",pattern="rwm_mcmc_1.Rda",recursive = T,full.names = T)

true.data.set <- str_extract(true.models,"Simulated_data_[0-9]+")
wrong.data.set <- str_extract(wrong.models,"Simulated_data_[0-9]+")

names(true.models) <- true.data.set
names(wrong.models) <- wrong.data.set

dic.truth <- lapply(true.models,getDIC) %>% bind_rows(.id="Data")
dic.wrong <- lapply(wrong.models,getDIC) %>% bind_rows(.id="Data")

dic.df <- dic.truth %>%
  inner_join(dic.wrong,by="Data",suffix=c("_True","_Wrong")) %>%
  mutate(Delta.DIC=DIC_Wrong - DIC_True)



```


```{r fig.width=5}

dic.hist <- ggplot(dic.df,aes(x=Delta.DIC)) +
  geom_bar() +
  scale_x_binned() +
  theme_cowplot() +
  xlab(expression(Delta*"DIC")) +
  ylab("Count") +
  ggtitle("Simulations suggest ability to distinguish which\ntrait drives the other")
dic.hist

dic.df.stationary <- dic.df %>%
  filter(Stationary_True == "Yes" & Stationary_Wrong == "Yes")

dic.hist.stat <- ggplot(dic.df.stationary,aes(x=Delta.DIC)) +
  geom_bar() +
  scale_x_binned() +
  xlab(expression(Delta*"DIC")) +
  theme_cowplot() +
  ggtitle("Truth: mRNA-driven model")
dic.hist.stat


dic.cutoff <- c(0,2,10,100,200,300)
names(dic.cutoff) <- dic.cutoff
delta.dic.score <- purrr::map_df(dic.cutoff,function(cutoff)
  {
   return((dic.df.stationary %>% filter(Delta.DIC > cutoff) %>% nrow()) / dic.df.stationary %>% filter(abs(Delta.DIC) > cutoff) %>% nrow()) 
}) %>% pivot_longer(everything(),names_to="Cutoff",values_to="Fraction_correct") %>%
  mutate(Cutoff = as.numeric(Cutoff))

dic.acc <- ggplot(delta.dic.score,aes(x=Cutoff,y=Fraction_correct)) +
  geom_point() +
  theme_cowplot() +
  xlab(expression(Delta*"DIC cutoff")) +
  ylab("Fraction of simulations correct model selected")
dic.acc

mrna.driven.dic.perf <- (dic.hist.stat | dic.acc)

overall.perf <- prot.driven.dic.perf / mrna.driven.dic.perf + plot_annotation(tag_levels = list(c("A","","B","")))
overall.perf

ggsave2("../Figures/dic_performance.pdf",overall.perf,width=11,height=10)


```




## Parameters differ across genes 

These represent cases where data is simulated under the assumption that parameters are not the same across genes, but follow a distribution.

### 100 genes w/o c fit 

```{r}
true.models <- list.files("Results/Truth_C_eq_0/100_loci/C_eq_0_fit/",pattern="Simulated_data_[0-9]+",full.names = T)
lapply(true.models,getPosteriorTable)
```


```{r fig.width=7,fig.height=7}

truth <- read_tsv("Simulated_data/true_parameters_used_for_parameters_diff_100_loci.tsv")
truth[,"H_Y"] <- truth[,"H_Y_Mean"]
truth[,"H_X"] <- truth[,"H_X_Mean"]
truth[,"Q"] <- truth[,"Q_Mean"]
truth[,"Sigma_Y"] <- truth[,"Sigma_Y_Mean"]
truth[,"Sigma_X"] <- truth[,"Sigma_X_Mean"]

truth <- truth %>%
  mutate(Simulation=paste0("Simulated_data_",1:nrow(truth))) %>%
  pivot_longer(-Simulation,names_to="Parameter",values_to="Truth")
  
directory <- "Results/Truth_C_eq_0/100_loci/C_eq_0_fit/"
model.fits <- list.files(directory,pattern="posterior_means.tsv", recursive = T)
names(model.fits) <- str_match(model.fits,"Simulated_data_[0-9]+")

post.mean.df <- lapply(model.fits,function(x)
  {
    read_tsv(file.path(directory,x),col_types =cols()) 
}
)%>% bind_rows(.id="Simulation")

post.mean.df <- post.mean.df %>%
  left_join(truth,by=c("Simulation","Parameter"))

post.mean.df.diff <- post.mean.df %>%
  filter(Parameter %in% c("H_Y","H_X","Q","C","Sigma_Y","Sigma_X"))
post.mean.df.diff <- post.mean.df.diff%>%
  mutate(Parameter = case_when(
    Parameter == "H_X" ~ "alpha[P]",
    Parameter == "H_Y" ~ "alpha[R]",
    Parameter == "Sigma_Y" ~ "tau[R]",
    Parameter == "Sigma_X" ~ "tau[P]",
    Parameter == "Q" ~ "a[R]",
    T ~ Parameter
  ))

p <- ggplot(post.mean.df.diff,aes(x=Truth,y=Mean)) +
  geom_point() +
  stat_cor(method="spearman",label.sep="\n") +
  geom_errorbar(aes(ymin=LB,ymax=UB)) +
  theme_cowplot()+
  geom_abline(intercept=0,slope=1,linetype="dashed") +
  theme(aspect.ratio=1) +
  facet_wrap(~Parameter,scales="free",labeller = "label_parsed")
p

```

### 100 genes w/c fits

```{r}
true.models <- list.files("Results/Truth_C_neq_0/100_loci/C_neq_0_fit/",pattern="Simulated_data_[0-9]+",full.names = T)
lapply(true.models,getPosteriorTable)
```
Models were simulated with a randomly chosen value of $c$. We then fit models with and without the $c$ parameter.


```{r fig.width=7,fig.height=7}

truth <- read_tsv("Simulated_data/true_parameters_used_for_parameters_diff_w_c_100_loci.tsv")
truth[,"H_Y"] <- truth[,"H_Y_Mean"]
truth[,"H_X"] <- truth[,"H_X_Mean"]
truth[,"Q"] <- truth[,"Q_Mean"]
truth[,"C"] <- truth[,"C_Mean"]
truth[,"Sigma_Y"] <- truth[,"Sigma_Y_Mean"]
truth[,"Sigma_X"] <- truth[,"Sigma_X_Mean"]

truth <- truth %>%
  mutate(Simulation=paste0("Simulated_data_",1:nrow(truth))) %>%
  pivot_longer(-Simulation,names_to="Parameter",values_to="Truth")
  
directory <- "Results/Truth_C_neq_0/100_loci/C_neq_0_fit/"
model.fits <- list.files(directory,pattern="posterior_means.tsv", recursive = T)
names(model.fits) <- str_match(model.fits,"Simulated_data_[0-9]+")

post.mean.df <- lapply(model.fits,function(x)
  {
    read_tsv(file.path(directory,x),col_types =cols()) 
}
)%>% bind_rows(.id="Simulation")

post.mean.df <- post.mean.df %>%
  left_join(truth,by=c("Simulation","Parameter"))

post.mean.df.diff <- post.mean.df %>%
  filter(Parameter %in% c("H_Y","H_X","Q","C","Sigma_Y","Sigma_X"))
post.mean.df.diff <- post.mean.df.diff%>%
  mutate(Parameter = case_when(
    Parameter == "H_X" ~ "alpha[R]",
    Parameter == "H_Y" ~ "alpha[P]",
    Parameter == "Sigma_Y" ~ "tau[P]",
    Parameter == "Sigma_X" ~ "tau[R]",
    Parameter == "Q" ~ "a[P]",
    Parameter == "C" ~ "c",
    T ~ Parameter
  ))

p <- ggplot(post.mean.df.diff,aes(x=Truth,y=Mean)) +
  geom_point() +
  stat_cor(method="spearman",label.sep="\n") +
  theme_cowplot()+
  geom_errorbar(aes(ymin=LB,ymax=UB)) +
  geom_abline(intercept=0,slope=1,linetype="dashed") +
  theme(aspect.ratio=1) +
  ylab("Posterior Mean") +
  facet_wrap(~Parameter,scales="free",labeller = "label_parsed")
p


ggsave2("../Figures/parameter_comparison_cneq_0.pdf",p,width=9,height=7)

```







