---
title: "Generate Simulated Data"
author: "Alex Cope"
date: '2023-05-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(PCMBase,lib.loc = "~/R_dev")
library(PCMBaseCpp)
library(tidyverse)


ba.tree <- ape::read.nexus("../Data/tree_11sp_noGpig.nex")
ba.tree <- phytools::force.ultrametric(ba.tree,method = "extend") # tree is not ultrametric due to apparent rounding error
# tip.heights <- ips::tipHeights(ba.tree)
# ba.tree <- pbtree(n = 50,scale = unname(tip.heights[1]))
# write.tree(ba.tree,"Simulated_data/Parameters_diff_same_prior_50_species/simulated_tree.nwk")
```

# Create OUOU model parameterization

```{r}
listParameterizationsOU <- PCMListParameterizations(structure(0.0, class="OU"))
listParameterizationsOU$X0 <- listParameterizationsOU$X0[1]
listParameterizationsOU$H <- listParameterizationsOU$H[12]
listParameterizationsOU$Sigma_x <- listParameterizationsOU$Sigma_x[2]
listParameterizationsOU$Sigmae_x <- listParameterizationsOU$Sigmae_x[c(5,7)]

dtParameterizations <- PCMTableParameterizations(
  structure(0.0, class="OU"), listParameterizations = listParameterizationsOU)

PCMGenerateParameterizations(structure(0.0, class="OU"), 
                             tableParameterizations = dtParameterizations[])

```


# Simulate datasets assuming parameters are truly shared across genes
Assume two traits $X$ and $Y$ where both traits are evolving under the an OU process.
The optimum of trait $Y$ at time $t$ $\theta(t)$ is a function of $X(t)$, i.e. $\theta(t) = \psi_Y + bX(t)$, where $\psi_Y$ is the optimum of $Y$ independent of the effects of $X$.
The optimum of $X$ will be denoted as $\psi_X$.
Assumes $\vec{\psi}$ is constant across the given tree.

The OU processes are assumed to be at stationarity, i.e. the root value of each trait $X(0)$ and $Y(0)$ is assumed to be the optimum for each trait.
This means $X(0) = \psi_X$ and $Y(0) = \psi_Y + bX(0) = \psi_Y + b\psi_X$.
This is based on equation D.1 from Bartoszeck et al. 2012.
Note that these derivation were based on the mvSlouch framework in which $X$ evolved according to a BM process.
Do we need to do something different?



```{r}
set.seed(100) # for reproducibility
num.sim <- 100
num.loci <- 100 # number of loci

parameter.matrix <- matrix(rep(0,num.sim*(num.loci*2+5)),nrow=num.sim,ncol = 2*num.loci + 5)
model.type <- "OU__Global_X0__Global_H__Theta__Diagonal_WithNonNegativeDiagonal_Sigma_x__Omitted_Sigmae_x"


ba.tree <- ape::read.nexus("../Data/tree_11sp_noGpig.nex")
ba.tree <- phytools::force.ultrametric(ba.tree,method = "extend") # tree is not ultrametric due to 
tip.height <- ips::tipHeights(ba.tree)[1]


sigma.x.mean <- -2.5
sigma.y.mean <- -2.5
alpha.x.mean <- log(log(2)/(tip.height/4))
alpha.y.mean <- log(log(2)/(tip.height/4))
b.mean <- 0.75

for (i in 1:num.sim)
{
  # Simulate the optima of the 2 traits across loci
  psi.x <- rnorm(n=num.loci,mean=0,sd=1)
  psi.y <- rnorm(n=num.loci,mean=0,sd=1)
  sigma.x <- rlnorm(n=1,meanlog = sigma.x.mean,sdlog = 1.5)
  sigma.y <- rlnorm(n=1,meanlog = sigma.y.mean,sdlog = 1.5)
  alpha.x <- rlnorm(n=1,meanlog = alpha.x.mean,sdlog = 1.5)
  alpha.y <- rlnorm(n=1,meanlog = alpha.y.mean,sdlog = 1.5)
  b <- rnorm(n=1,b.mean,0.25)
  
  ## Get root state
  x0 <- psi.x 
  y0 <- psi.y + b*x0

  data.pcm <- vector(mode="list",length=num.loci)
  true.param <- c(alpha.y,alpha.x,b,sigma.y,sigma.x)
  names(true.param) <- c("H_Y","H_X","Q","Sigma_Y","Sigma_X")
  for(j in 1:num.loci)
  {
    true.model <- PCM(model=model.type,k=2)
    param <- c(y0[j],
                    x0[j],
                    alpha.y,
                    0,
                    -b*alpha.y,
                    alpha.x,
                    y0[j],
                    x0[j],
                    sigma.y,
                    sigma.x)
    PCMParamLoadOrStore(true.model, param, offset=0, load=T)
    
    psi.param <- c(psi.y[j],psi.x[j])
    names(psi.param) <- paste0(c("Psi_Y_","Psi_X_"),j)
    true.param <- c(true.param,psi.param)
    data.pcm[[j]] <- PCMSim(tree = ba.tree,model = true.model,X0 = true.model$X0)
  
  }
  parameter.matrix[i,] <- true.param
  names(data.pcm) <- paste0("Gene_",rep(1:num.loci))
  data.df <- purrr::map(data.pcm,~.x %>% 
                          t() %>% 
                          as.data.frame() %>% 
                          rownames_to_column("Species")) %>% 
                        bind_rows(.id="Locus") %>%
             filter(Species %in% ba.tree$tip.label)
  write_tsv(data.df, file.path("Simulated_data/Parameters_same_100_loci/",paste0("simulated_data_",i)))
}

colnames(parameter.matrix) <-c("H_Y","H_X","Q","Sigma_Y","Sigma_X",paste0(c("Psi_X_","Psi_Y_"),rep(1:num.loci,each=2)))


write_tsv(as.data.frame(parameter.matrix),"Simulated_data/true_parameters_used_for_parameters_same_sim_100_loci.tsv")
```




# Simulate datasets assuming parameters are different across genes


## Without C

### 50 genes
```{r}
set.seed(50) # for reproducibility
num.sim <- 100
num.loci <- 50 # number of loci

ba.tree <- ape::read.nexus("../Data/tree_11sp_noGpig.nex")
ba.tree <- phytools::force.ultrametric(ba.tree,method = "extend") # tree is not ultrametric due to 
tip.height <- ips::tipHeights(ba.tree)[1]
min.edge <- min(ba.tree$edge.length)
parameter.matrix <- matrix(rep(0,num.sim*(num.loci*7+10)),nrow=num.sim,ncol = 7*num.loci + 10)
model.type <- "OU__Global_X0__Global_H__Theta__Diagonal_WithNonNegativeDiagonal_Sigma_x__Omitted_Sigmae_x"

for (i in 1:num.sim)
{
  # Simulate the optima of the 2 traits across loci
  psi.x <- rnorm(n=num.loci,mean=0,sd=1)
  psi.y <- rnorm(n=num.loci,mean=0,sd=1)
  
  sigma.x.mean <- runif(n=1,min=-3,max=-2)
  sigma.y.mean <- runif(n=1,min=-3,max=-2)
  alpha.x.mean <- runif(n=1,min=log(log(2)/(tip.height/4)),max=log(log(2)/(min.edge*2)))
  alpha.y.mean <- runif(n=1,min=log(log(2)/(tip.height/4)),max=log(log(2)/(min.edge*2)))
  b.mean <- runif(n=1,min=0,max=1)
  
  data.pcm <- vector(mode="list",length=num.loci)
  true.param <- c(alpha.y.mean,log(0.5),alpha.x.mean,log(0.5),b.mean,log(0.25),sigma.y.mean,log(0.5),sigma.x.mean,log(0.5))
  names(true.param) <- c("H_Y_Mean","H_Y_Std","H_X_Mean","H_X_Std","Q_Mean","Q_Std","Sigma_Y_Mean","Sigma_Y_Std","Sigma_X_Mean","Sigma_X_Std")
  for(j in 1:num.loci)
  {
    sigma.x <- rlnorm(n=1,meanlog = sigma.x.mean,sdlog = 0.5)
    sigma.y <- rlnorm(n=1,meanlog = sigma.y.mean,sdlog = 0.5)
    alpha.x <- rlnorm(n=1,meanlog = alpha.x.mean,sdlog = 0.5)
    alpha.y <- rlnorm(n=1,meanlog = alpha.y.mean,sdlog = 0.5)
    b <- rnorm(n=1,mean=b.mean,sd=0.25)
    ## Get root state
    x0 <- psi.x[j]
    y0 <- psi.y[j] + b*x0
    true.model <- PCM(model=model.type,k=2)
    param <- c(y0,
                    x0,
                    alpha.y,
                    0,
                    -b*alpha.y,
                    alpha.x,
                    y0,
                    x0,
                    sigma.y,
                    sigma.x)
    PCMParamLoadOrStore(true.model, param, offset=0, load=T)
    
    gene.param <- c(psi.y[j],psi.x[j],log(alpha.y),log(alpha.x),b,log(sigma.y),log(sigma.x))
    names(gene.param) <- paste0(c("Psi_Y_","Psi_X_","H_Y_","H_X_","Q_","Sigma_Y_","Sigma_X_"),j)
    true.param <- c(true.param,gene.param)
    data.pcm[[j]] <- PCMSim(tree = ba.tree,model = true.model,X0 = true.model$X0)
  
  }
  parameter.matrix[i,] <- true.param
  names(data.pcm) <- paste0("Gene_",rep(1:num.loci))
  data.df <- purrr::map(data.pcm,~.x %>% 
                          t() %>% 
                          as.data.frame() %>% 
                          rownames_to_column("Species")) %>% 
                        bind_rows(.id="Locus") %>%
             filter(Species %in% ba.tree$tip.label)
  write_tsv(data.df, file.path("Simulated_data/Parameters_diff_50_loci_0.5_sdlog/",paste0("simulated_data_",i)))
}

colnames(parameter.matrix) <-c("H_Y_Mean","H_Y_Std","H_X_Mean","H_X_Std","Q_Mean","Q_Std","Sigma_Y_Mean","Sigma_Y_Std","Sigma_X_Mean","Sigma_X_Std",paste0(c("Psi_Y_","Psi_X_","H_Y_","H_X_","Q_","Sigma_Y_","Sigma_X_"),rep(1:num.loci,each=7)))


write_tsv(as.data.frame(parameter.matrix),"Simulated_data/true_parameters_used_for_parameters_diff_50_loci.tsv")
```



```{r}
set.seed(50) # for reproducibility
num.sim <- 100
num.loci <- 100 # number of loci

ba.tree <- ape::read.nexus("../Data/tree_11sp_noGpig.nex")
ba.tree <- phytools::force.ultrametric(ba.tree,method = "extend") # tree is not ultrametric due to 
tip.height <- ips::tipHeights(ba.tree)[1]
min.edge <- min(ba.tree$edge.length)
parameter.matrix <- matrix(rep(0,num.sim*(num.loci*7+10)),nrow=num.sim,ncol = 7*num.loci + 10)
model.type <- "OU__Global_X0__Global_H__Theta__Diagonal_WithNonNegativeDiagonal_Sigma_x__Omitted_Sigmae_x"

for (i in 1:num.sim)
{
  # Simulate the optima of the 2 traits across loci
  psi.x <- rnorm(n=num.loci,mean=0,sd=1)
  psi.y <- rnorm(n=num.loci,mean=0,sd=1)
  
  sigma.x.mean <- runif(n=1,min=-3,max=-2)
  sigma.y.mean <- runif(n=1,min=-3,max=-2)
  alpha.x.mean <- runif(n=1,min=log(log(2)/(tip.height/4)),max=log(log(2)/(min.edge*2)))
  alpha.y.mean <- runif(n=1,min=log(log(2)/(tip.height/4)),max=log(log(2)/(min.edge*2)))
  b.mean <- runif(n=1,min=0,max=1)
  
  data.pcm <- vector(mode="list",length=num.loci)
  true.param <- c(alpha.y.mean,log(0.5),alpha.x.mean,log(0.5),b.mean,log(0.25),sigma.y.mean,log(0.5),sigma.x.mean,log(0.5))
  names(true.param) <- c("H_Y_Mean","H_Y_Std","H_X_Mean","H_X_Std","Q_Mean","Q_Std","Sigma_Y_Mean","Sigma_Y_Std","Sigma_X_Mean","Sigma_X_Std")
  for(j in 1:num.loci)
  {
    sigma.x <- rlnorm(n=1,meanlog = sigma.x.mean,sdlog = 0.5)
    sigma.y <- rlnorm(n=1,meanlog = sigma.y.mean,sdlog = 0.5)
    alpha.x <- rlnorm(n=1,meanlog = alpha.x.mean,sdlog = 0.5)
    alpha.y <- rlnorm(n=1,meanlog = alpha.y.mean,sdlog = 0.5)
    b <- rnorm(n=1,mean=b.mean,sd=0.25)
    ## Get root state
    x0 <- psi.x[j]
    y0 <- psi.y[j] + b*x0
    true.model <- PCM(model=model.type,k=2)
    param <- c(y0,
                    x0,
                    alpha.y,
                    0,
                    -b*alpha.y,
                    alpha.x,
                    y0,
                    x0,
                    sigma.y,
                    sigma.x)
    PCMParamLoadOrStore(true.model, param, offset=0, load=T)
    
    gene.param <- c(psi.y[j],psi.x[j],log(alpha.y),log(alpha.x),b,log(sigma.y),log(sigma.x))
    names(gene.param) <- paste0(c("Psi_Y_","Psi_X_","H_Y_","H_X_","Q_","Sigma_Y_","Sigma_X_"),j)
    true.param <- c(true.param,gene.param)
    data.pcm[[j]] <- PCMSim(tree = ba.tree,model = true.model,X0 = true.model$X0)
  
  }
  parameter.matrix[i,] <- true.param
  names(data.pcm) <- paste0("Gene_",rep(1:num.loci))
  data.df <- purrr::map(data.pcm,~.x %>% 
                          t() %>% 
                          as.data.frame() %>% 
                          rownames_to_column("Species")) %>% 
                        bind_rows(.id="Locus") %>%
             filter(Species %in% ba.tree$tip.label)
  write_tsv(data.df, file.path("Simulated_data/Parameters_diff_100_loci_0.5_sdlog/",paste0("simulated_data_",i)))
}

colnames(parameter.matrix) <-c("H_Y_Mean","H_Y_Std","H_X_Mean","H_X_Std","Q_Mean","Q_Std","Sigma_Y_Mean","Sigma_Y_Std","Sigma_X_Mean","Sigma_X_Std",paste0(c("Psi_Y_","Psi_X_","H_Y_","H_X_","Q_","Sigma_Y_","Sigma_X_"),rep(1:num.loci,each=7)))


write_tsv(as.data.frame(parameter.matrix),"Simulated_data/true_parameters_used_for_parameters_diff_100_loci.tsv")
```


## With C

### 50 genes
```{r}
set.seed(50) # for reproducibility
num.sim <- 100
num.loci <- 50 # number of loci

ba.tree <- ape::read.nexus("../Data/tree_11sp_noGpig.nex")
ba.tree <- phytools::force.ultrametric(ba.tree,method = "extend") # tree is not ultrametric due to 
tip.height <- ips::tipHeights(ba.tree)[1]
min.edge <- min(ba.tree$edge.length)
parameter.matrix <- matrix(rep(0,num.sim*(num.loci*8+12)),nrow=num.sim,ncol = 8*num.loci + 12)
model.type <- "OU__Global_X0__Global_H__Theta__UpperTriangularWithDiagonal_WithNonNegativeDiagonal_Sigma_x__Omitted_Sigmae_x"

for (i in 1:num.sim)
{
  # Simulate the optima of the 2 traits across loci
  psi.x <- rnorm(n=num.loci,mean=0,sd=1)
  psi.y <- rnorm(n=num.loci,mean=0,sd=1)
  
  sigma.x.mean <- runif(n=1,min=-3,max=-2)
  sigma.y.mean <- runif(n=1,min=-3,max=-2)
  alpha.x.mean <- runif(n=1,min=log(log(2)/(tip.height/4)),max=log(log(2)/(min.edge*2)))
  alpha.y.mean <- runif(n=1,min=log(log(2)/(tip.height/4)),max=log(log(2)/(min.edge*2)))
  b.mean <- runif(n=1,min=0,max=1)
  c.mean <- runif(n=1,min=-1,max=1)
  
  data.pcm <- vector(mode="list",length=num.loci)
  true.param <- c(alpha.y.mean,log(0.5),alpha.x.mean,log(0.5),b.mean,log(0.25),c.mean,log(0.25),sigma.y.mean,log(0.5),sigma.x.mean,log(0.5))
  names(true.param) <- c("H_Y_Mean","H_Y_Std","H_X_Mean","H_X_Std","Q_Mean","Q_Std","C_Mean","C_Std","Sigma_Y_Mean","Sigma_Y_Std","Sigma_X_Mean","Sigma_X_Std")
  for(j in 1:num.loci)
  {
    sigma.x <- rlnorm(n=1,meanlog = sigma.x.mean,sdlog = 0.5)
    sigma.y <- rlnorm(n=1,meanlog = sigma.y.mean,sdlog = 0.5)
    alpha.x <- rlnorm(n=1,meanlog = alpha.x.mean,sdlog = 0.5)
    alpha.y <- rlnorm(n=1,meanlog = alpha.y.mean,sdlog = 0.5)
    b <- rnorm(n=1,mean=b.mean,sd=0.25)
    c.val <- rnorm(n=1,mean=c.mean,sd=0.25)
    ## Get root state
    x0 <- psi.x[j]
    y0 <- psi.y[j] + (b - (alpha.x*c.val)/alpha.y)*x0
    true.model <- PCM(model=model.type,k=2)
    param <- c(y0,
                    x0,
                    alpha.y,
                    0,
                    alpha.x * c.val -b * alpha.y,
                    alpha.x,
                    y0,
                    x0,
                    sigma.y,
                    c.val*sigma.x,
                    sigma.x)
    PCMParamLoadOrStore(true.model, param, offset=0, load=T)
    
    gene.param <- c(psi.y[j],psi.x[j],log(alpha.y),log(alpha.x),b,c.val,log(sigma.y),log(sigma.x))
    names(gene.param) <- paste0(c("Psi_Y_","Psi_X_","H_Y_","H_X_","Q_","C_","Sigma_Y_","Sigma_X_"),j)
    true.param <- c(true.param,gene.param)
    data.pcm[[j]] <- PCMSim(tree = ba.tree,model = true.model,X0 = true.model$X0)
  
  }
  parameter.matrix[i,] <- true.param
  names(data.pcm) <- paste0("Gene_",rep(1:num.loci))
  data.df <- purrr::map(data.pcm,~.x %>% 
                          t() %>% 
                          as.data.frame() %>% 
                          rownames_to_column("Species")) %>% 
                        bind_rows(.id="Locus") %>%
             filter(Species %in% ba.tree$tip.label)
  write_tsv(data.df, file.path("Simulated_data/Parameters_diff_w_c_50_loci_0.5_sdlog/",paste0("simulated_data_",i)))
}

colnames(parameter.matrix) <-c("H_Y_Mean","H_Y_Std","H_X_Mean","H_X_Std","Q_Mean","Q_Std","C_Mean","C_Std","Sigma_Y_Mean","Sigma_Y_Std","Sigma_X_Mean","Sigma_X_Std",paste0(c("Psi_Y_","Psi_X_","H_Y_","H_X_","Q_","C_","Sigma_Y_","Sigma_X_"),rep(1:num.loci,each=8)))


write_tsv(as.data.frame(parameter.matrix),"Simulated_data/true_parameters_used_for_parameters_diff_w_c_50_loci.tsv")
```


### 100 genes
```{r}
set.seed(50) # for reproducibility
num.sim <- 100
num.loci <- 100 # number of loci

ba.tree <- ape::read.nexus("../Data/tree_11sp_noGpig.nex")
ba.tree <- phytools::force.ultrametric(ba.tree,method = "extend") # tree is not ultrametric due to 
tip.height <- ips::tipHeights(ba.tree)[1]
min.edge <- min(ba.tree$edge.length)
parameter.matrix <- matrix(rep(0,num.sim*(num.loci*8+12)),nrow=num.sim,ncol = 8*num.loci + 12)
model.type <- "OU__Global_X0__Global_H__Theta__UpperTriangularWithDiagonal_WithNonNegativeDiagonal_Sigma_x__Omitted_Sigmae_x"

for (i in 1:num.sim)
{
  # Simulate the optima of the 2 traits across loci
  psi.x <- rnorm(n=num.loci,mean=0,sd=1)
  psi.y <- rnorm(n=num.loci,mean=0,sd=1)
  
  sigma.x.mean <- runif(n=1,min=-3,max=-2)
  sigma.y.mean <- runif(n=1,min=-3,max=-2)
  alpha.x.mean <- runif(n=1,min=log(log(2)/(tip.height/4)),max=log(log(2)/(min.edge*2)))
  alpha.y.mean <- runif(n=1,min=log(log(2)/(tip.height/4)),max=log(log(2)/(min.edge*2)))
  b.mean <- runif(n=1,min=0,max=1)
  c.mean <- runif(n=1,min=-1,max=1)
  
  data.pcm <- vector(mode="list",length=num.loci)
  true.param <- c(alpha.y.mean,log(0.5),alpha.x.mean,log(0.5),b.mean,log(0.25),c.mean,log(0.25),sigma.y.mean,log(0.5),sigma.x.mean,log(0.5))
  names(true.param) <- c("H_Y_Mean","H_Y_Std","H_X_Mean","H_X_Std","Q_Mean","Q_Std","C_Mean","C_Std","Sigma_Y_Mean","Sigma_Y_Std","Sigma_X_Mean","Sigma_X_Std")
  for(j in 1:num.loci)
  {
    sigma.x <- rlnorm(n=1,meanlog = sigma.x.mean,sdlog = 0.5)
    sigma.y <- rlnorm(n=1,meanlog = sigma.y.mean,sdlog = 0.5)
    alpha.x <- rlnorm(n=1,meanlog = alpha.x.mean,sdlog = 0.5)
    alpha.y <- rlnorm(n=1,meanlog = alpha.y.mean,sdlog = 0.5)
    b <- rnorm(n=1,mean=b.mean,sd=0.25)
    c.val <- rnorm(n=1,mean=c.mean,sd=0.25)
    ## Get root state
    x0 <- psi.x[j]
    y0 <- psi.y[j] + b*x0
    true.model <- PCM(model=model.type,k=2)
    param <- c(y0,
                    x0,
                    alpha.y,
                    0,
                    alpha.x * c.val - b*alpha.y,
                    alpha.x,
                    y0,
                    x0,
                    sigma.y,
                    c.val*sigma.x,
                    sigma.x)
    PCMParamLoadOrStore(true.model, param, offset=0, load=T)
    
    gene.param <- c(psi.y[j],psi.x[j],log(alpha.y),log(alpha.x),b,c.val,log(sigma.y),log(sigma.x))
    names(gene.param) <- paste0(c("Psi_Y_","Psi_X_","H_Y_","H_X_","Q_","C_","Sigma_Y_","Sigma_X_"),j)
    true.param <- c(true.param,gene.param)
    data.pcm[[j]] <- PCMSim(tree = ba.tree,model = true.model,X0 = true.model$X0)
  
  }
  parameter.matrix[i,] <- true.param
  names(data.pcm) <- paste0("Gene_",rep(1:num.loci))
  data.df <- purrr::map(data.pcm,~.x %>% 
                          t() %>% 
                          as.data.frame() %>% 
                          rownames_to_column("Species")) %>% 
                        bind_rows(.id="Locus") %>%
             filter(Species %in% ba.tree$tip.label)
  write_tsv(data.df, file.path("Simulated_data/Parameters_diff_w_c_100_loci_0.5_sdlog/",paste0("simulated_data_",i)))
}

colnames(parameter.matrix) <-c("H_Y_Mean","H_Y_Std","H_X_Mean","H_X_Std","Q_Mean","Q_Std","C_Mean","C_Std","Sigma_Y_Mean","Sigma_Y_Std","Sigma_X_Mean","Sigma_X_Std",paste0(c("Psi_Y_","Psi_X_","H_Y_","H_X_","Q_","C_","Sigma_Y_","Sigma_X_"),rep(1:num.loci,each=8)))


write_tsv(as.data.frame(parameter.matrix),"Simulated_data/true_parameters_used_for_parameters_diff_w_c_100_loci.tsv")
```

### 164 genes
```{r}
set.seed(50) # for reproducibility
num.sim <- 100
num.loci <- 164 # number of loci

ba.tree <- ape::read.nexus("../Data/tree_11sp_noGpig.nex")
ba.tree <- phytools::force.ultrametric(ba.tree,method = "extend") # tree is not ultrametric due to 
tip.height <- ips::tipHeights(ba.tree)[1]
min.edge <- min(ba.tree$edge.length)
parameter.matrix <- matrix(rep(0,num.sim*(num.loci*8+12)),nrow=num.sim,ncol = 8*num.loci + 12)
model.type <- "OU__Global_X0__Global_H__Theta__UpperTriangularWithDiagonal_WithNonNegativeDiagonal_Sigma_x__Omitted_Sigmae_x"

for (i in 1:num.sim)
{
  # Simulate the optima of the 2 traits across loci
  psi.x <- rnorm(n=num.loci,mean=0,sd=1)
  psi.y <- rnorm(n=num.loci,mean=0,sd=1)
  
  sigma.x.mean <- runif(n=1,min=-3,max=-2)
  sigma.y.mean <- runif(n=1,min=-3,max=-2)
  alpha.x.mean <- runif(n=1,min=log(log(2)/(tip.height/4)),max=log(log(2)/(min.edge*2)))
  alpha.y.mean <- runif(n=1,min=log(log(2)/(tip.height/4)),max=log(log(2)/(min.edge*2)))
  b.mean <- runif(n=1,min=0,max=1)
  c.mean <- runif(n=1,min=-1,max=1)
  
  data.pcm <- vector(mode="list",length=num.loci)
  true.param <- c(alpha.y.mean,log(0.5),alpha.x.mean,log(0.5),b.mean,log(0.25),c.mean,log(0.25),sigma.y.mean,log(0.5),sigma.x.mean,log(0.5))
  names(true.param) <- c("H_Y_Mean","H_Y_Std","H_X_Mean","H_X_Std","Q_Mean","Q_Std","C_Mean","C_Std","Sigma_Y_Mean","Sigma_Y_Std","Sigma_X_Mean","Sigma_X_Std")
  for(j in 1:num.loci)
  {
    sigma.x <- rlnorm(n=1,meanlog = sigma.x.mean,sdlog = 0.5)
    sigma.y <- rlnorm(n=1,meanlog = sigma.y.mean,sdlog = 0.5)
    alpha.x <- rlnorm(n=1,meanlog = alpha.x.mean,sdlog = 0.5)
    alpha.y <- rlnorm(n=1,meanlog = alpha.y.mean,sdlog = 0.5)
    b <- rnorm(n=1,mean=b.mean,sd=0.25)
    c.val <- rnorm(n=1,mean=c.mean,sd=0.25)
    ## Get root state
    x0 <- psi.x[j]
    y0 <- psi.y[j] + b*x0
    true.model <- PCM(model=model.type,k=2)
    param <- c(y0,
                    x0,
                    alpha.y,
                    0,
                    alpha.x * c.val - b*alpha.y,
                    alpha.x,
                    y0,
                    x0,
                    sigma.y,
                    c.val*sigma.x,
                    sigma.x)
    PCMParamLoadOrStore(true.model, param, offset=0, load=T)
    
    gene.param <- c(psi.y[j],psi.x[j],log(alpha.y),log(alpha.x),b,c.val,log(sigma.y),log(sigma.x))
    names(gene.param) <- paste0(c("Psi_Y_","Psi_X_","H_Y_","H_X_","Q_","C_","Sigma_Y_","Sigma_X_"),j)
    true.param <- c(true.param,gene.param)
    data.pcm[[j]] <- PCMSim(tree = ba.tree,model = true.model,X0 = true.model$X0)
  
  }
  parameter.matrix[i,] <- true.param
  names(data.pcm) <- paste0("Gene_",rep(1:num.loci))
  data.df <- purrr::map(data.pcm,~.x %>% 
                          t() %>% 
                          as.data.frame() %>% 
                          rownames_to_column("Species")) %>% 
                        bind_rows(.id="Locus") %>%
             filter(Species %in% ba.tree$tip.label)
  write_tsv(data.df, file.path("Simulated_data/Parameters_diff_w_c_164_loci_0.5_sdlog/",paste0("simulated_data_",i)))
}

colnames(parameter.matrix) <-c("H_Y_Mean","H_Y_Std","H_X_Mean","H_X_Std","Q_Mean","Q_Std","C_Mean","C_Std","Sigma_Y_Mean","Sigma_Y_Std","Sigma_X_Mean","Sigma_X_Std",paste0(c("Psi_Y_","Psi_X_","H_Y_","H_X_","Q_","C_","Sigma_Y_","Sigma_X_"),rep(1:num.loci,each=8)))


write_tsv(as.data.frame(parameter.matrix),"Simulated_data/true_parameters_used_for_parameters_diff_w_c_164_loci.tsv")
```


